---
title: "Problem Set 2"
author: "Marc Eskew"
date: "4/6/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggforce)
```

## Problem 1

We first calculate $Cov(e^\frac{T}{2},e^\frac{T}{3})$ using integration:

$$
\begin{aligned}
E[X] &= \int_{0}^{\infty} e^\frac{t}{2}e^{-t} dt \\
&= 2 \\
E[Y] &= \int_{0}^{\infty} e^\frac{t}{3}e^{-t} dt \\
&= 1.5 \\
Cov(e^\frac{T}{2},e^\frac{T}{3}) &= \int_{0}^{\infty}(x - E[X])(y - E[Y]f_T(t) dt \\
&= \int_{0}^{\infty}(e^\frac{t}{2}-2)(e^\frac{t}{3}-1.5)(e^{-t}) dt \\
&= 3\\
\end{aligned}
$$

```{r problem_1}
reps <- c(100,1000,10000,100000)

for(i in reps) {

  t_vec <- rexp(i)
  
  df <- data.frame(t = t_vec) %>%
    mutate(t_1 = exp(t/2),
           t_2 = exp(t/3), t_3 = t_1*t_2)
  
  cov_est <- mean(df$t_3) - mean(df$t_1)*mean(df$t_2)
  
  print(paste(i,":",cov_est))

}
```

Based on these experiments, I do not think this converges at a speed proportional to $\frac{1}{\sqrt{N}}$.  This does not contradict our discussion, but what fails here is the consistency of determining the variance of one replication of the simulation.


## Problem 2


```{r multi_sample}
x <- 2*runif(10000,0,1)
y <- 2*runif(10000,0,1)

df <- data.frame(x=x, y=y) %>%
  mutate(i = ifelse(y<=x,1,0) 
  ) %>%
  mutate(g = sin(x*y)*i)

est <- 4*mean(df$g)
print(est)
```

The MC estimation over this simulation run was **`r est`**.  This is near the solution of ~1.05 determined through calculus.


## Problem 3

With a $\lambda(t)$ that increases linearly with time, I would expect the component modeled as an exponential would last longer.

Using the inversion method to sample $T$ with $\lambda(t) = 2t + 1$:

$$
\begin{aligned}
\Lambda(t) &= \int_0^t2t+1\,dt \\
&= t^2 + t \\
F_Y(y) &= u \\
u &= 1-e^{-t^2-t} \\
\frac{1}{1-u} &= e^{t^2+t} \\
log(\frac{1}{1-u}) &= t^2 + t \\
log(\frac{1}{1-u}) + \frac{1}{4} &= (t+\frac{1}{2})^2 \\
\sqrt{log(\frac{1}{1-u}) + \frac{1}{4}} &= t+\frac{1}{2} \\
\sqrt{log(\frac{1}{1-u}) + \frac{1}{4}} - \frac{1}{2} &= t
\end{aligned}
$$

With the cumulative function inverted, we can sample from a $U(0,1)$ distribution to represent $T$.

```{r r_plot}
exp_life <- rexp(10000) # Exponential comparison

T_life <- data.frame(u = runif(10000)) %>% # Sample from U(0,1)
  mutate(t = sqrt(.25 + log(1/(1-u)))-.5) # Apply inverse function

ggplot(T_life) +
  geom_density(aes(x = t, color = "T")) +
  geom_density(data = data.frame(x=exp_life),aes(x=x,color = "Exp(1)")) +
  ggsci::scale_color_aaas() +
  ggtitle("PDF Comparison")

```

The mean lifetime when modeled by $T$ is **`r mean(T_life$t)`**, which is lower than the exponential model.  

## Problem 4

### a. $\mathbb{E}[N^2]$

```{r question4a}
df_out <- data.frame()

for(i in 1:10000) {
  n <- -1
  p <- 1
  
  while(p > exp(-9)) {
    p <- p*runif(1)^3
    n <- n+1
  }
  
  df_out <- rbind(df_out,n)
}

colnames(df_out) <- "N"
mean(df_out$N^2)

```

### b.$\mathbb{P}[N=i]$


```{r question4b}
for(i in 0:6) {
  df_out1 <- df_out %>%
    filter(N == i)
  
  print(paste(i,":",nrow(df_out1)/nrow(df_out)))
}
```

